---
title: "scRNAseq_hermvsmale_analysis"
format: html
editor: visual
---

## Purpose

The purpose of this code is to analyze/visualize the scRNAseq data from Meital, so that we can better select NPs fo interest to work with in the future

Importing libraries

```{r}
#| label: loading libraries
#| output: false
library(tidyverse)
```

Reading in files

```{r}
path2wd <- "/Users/tammyray/Desktop/hart_code"
setwd(paste0(path2wd,"/data_sheets"))
dvb_mean_expression <- read_csv("ScRNA_DVB_Meital_Mean_expression.csv")
colnames(dvb_mean_expression)[1] <- "gene"
avl <- read_csv("ScRNA_AVL_Meital.csv")
dvb <- read_csv("ScRNA_DVB_Meital.csv")
dvc <- read_csv("ScRNA_DVC_Meital.csv")
```

Tidying the dataframes

```{r}
colnames(dvb_mean_expression)[2] <- "Herm_mean"
colnames(dvb_mean_expression)[3] <- "Male_mean"

dvb_mean_expression <- dvb_mean_expression %>%
  mutate(Herm_mean = round(Herm_mean, digits=4)) %>%
  mutate(Male_mean = round(Male_mean, digits=4)) %>%
  mutate(Male_minus_Herm = Male_mean - Herm_mean)

```

```{r}
#Now I want to make a new dataframe for DVB data, that basically combines my dvb_mean_expression database with logfoldchange and adjusted p value

#First, i will make a new dataframe from dvb that only includes the things I want to add to my new database
dvb_stripped <- dvb %>%
  select(-c("p_val", "pct.1", "pct.2", "Comp", "Partition", "Downsample", "Male_threshold", "Herm_threshold", "Male_cells", "Herm_cells"))
colnames(dvb_stripped)[3] <- "gene"

dvb_all_info <- dvb_mean_expression %>%
  left_join(dvb_stripped, by = "gene")
```

```{r}
# Hmmm, now I want to filter out all non-significant adjusted p values...
sex_shared_cells <- all_cells %>%
  filter(ID == "sex-shared")
write_csv(sex_shared_cells, "Sex_shared_cells.csv")
```

## How the data looks

I'm curious, how many cells do we need to be able to say something is significant at alpha=0.05 with 80% confidence? If we are looking at large effect sizes, cohen's d=0.8, then we would need around 26 cells... I remember this math from my quant neuro presentation, but here's the code to do that:

```{r}
#| label: Loading necessary libraries
#| output: false
library(pwr)
```

```{r}
# Defining my paramenters for the power analysis 
effect_size <- 0.8  # Cohen's d (large effect size)
alpha <- 0.05       # Significance level
power <- 0.8        # Desired power (80%)

# Calculate sample size needed for the given effect size, alpha, and power
result <- pwr.t.test(d = effect_size, sig.level = alpha, power = power, type = "two.sample")
print(result)
```

## ALL CELLS IN C ELEGAN FOR MALE AND HERMAPHRODITE

I downloaded and then organized data from Supplementary Information 4 from [2019 Whole-animal connectomes of both Caenorhabditis elegans sexes](https://www.nature.com/articles/s41586-019-1352-7#Sec27).

**Description:**\
"Cell Lists These sheets list all of the cells in the *C. elegans* adult nervous system along with end organs as they appear in the connectivity matrices."

**The data contains the following variables**

-   cell_name - the name of each cell in either sex C. elegans

-   cell_type - if it is a sensory neuron, interneuron, muscle, etc.

-   notes_cellCategory

-   ID - if the cell is male-specific, herm-specific, or faoun in both, or in the pharynx

-   notes_sensilla.modality - sensory modality information for SOME of the sex-sahred neurons, also some random notes "interneuron in White et al., 1986"

```{r}
setwd("/Users/tammyray/Desktop/hart_code/data_sheets")
all_cells <- read_csv("ALL_adult_c.elegan_cells_Emmons2019_supp_info_4")
```

## A New day: Making my giant database

I want to make a database that contains all connectome information(synaptic chemical, synaptic electrical, monoaminergic, and neuropeptidergic).

I am writing notes on this in my notion.

First, I want to input the synaptic electrical and chemical databases I have for male and hermaphrodite into r. I want to QC for duplicate entries in the pre and post cell (when I looked at one fo the sheets, I thought I saw the exact same thing for one of them

```{r}
setwd(paste0(path2wd,"/data_sheets"))
male_synaptic_connectome_raw <- read_csv("Male_synaptic_connectome.csv")
```
